### 2023-09-19 04_manuscript/test_grateful.Rmd

---
title: "Testing the grateful package"
output: pdf_document
date: "2022-09-22"
bibliography: ["test.bib", "grateful-refs.bib"]
csl: ecology-letters.csl
#csl: "https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

Let's load libraries.

```{r loadlib, message = FALSE}
library(janitor)
library(tidyverse)
library(palmerpenguins)
library(grateful)
```

let's now create a BibTex file that includes citations for these libraries.

```{r}
get_pkgs_info()
```

And that function automatically produces a .bib file called "grateful-refs.bib". So we'll add that to our YAML.

## Tables and Plots

I'll use the [@mouillotDimensionalityStructureSpecies2021] a lot in my writing. I want to add another citation here [@denelleDistinguishingSignaturesLocal2019].

And here i'd like to use the `palmerpenguins` package

```{r crosstab}

penguins %>%
  tabyl(species, island) %>%
  adorn_totals(where = c("row", "col"))

```

Now a figure

```{r}
penguins %>% 
  filter(sex == "male") %>%
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin() +
  geom_jitter(size = 2, shape = 1, width = 0.1) +
  xlab("Species") +
  ylab("Body mass (g)") +
  theme_bw()
```

```{r results = 'asis'}
cite_packages(output = "paragraph")
```

\newpage

## References