### 2023-09-28 05_preregistration/OSF_preregistration.tex
Update homework

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Succession forestière après feu: Où la forêt en est-elle rendue 100 ans plus tard?},
  pdfauthor={Ariane Roberge},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Succession forestière après feu: Où la forêt en est-elle rendue
100 ans plus tard?}
\author{Ariane Roberge}
\date{2023-09-19}

\begin{document}
\maketitle

\hypertarget{study-information}{%
\subsection{Study Information}\label{study-information}}

\hypertarget{title}{%
\subsubsection{1. Title}\label{title}}

Succession forestière après feu : Où la forêt en est-elle rendue 100 ans
plus tard?

\hypertarget{authorship}{%
\subsubsection{2. Authorship}\label{authorship}}

Ariane Roberge\textsuperscript{1} Etienne Laliberté\textsuperscript{1}

\textsuperscript{1} Département de Sciences Biologiques, Institut de
Recherche en Biologie Végétale, Université de Montréal, 4101 Sherbrooke
Est, Montréal, QC H1X 2B2, Canada.

\hypertarget{description}{%
\subsubsection{3. Description}\label{description}}

Enter a description of your study here.

\hypertarget{hypotheses}{%
\subsubsection{4. Hypotheses}\label{hypotheses}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  List any \emph{a priori} hypotheses here.
\end{enumerate}

\hypertarget{design-plan}{%
\subsection{Design Plan}\label{design-plan}}

\hypertarget{study-type}{%
\subsubsection{5. Study Type}\label{study-type}}

\begin{itemize}
\tightlist
\item
  Observational Study - Data is collected from study subjects that are
  not randomly assigned to a treatment. This includes surveys, natural
  experiments, and regression discontinuity designs.
\end{itemize}

\hypertarget{blinding}{%
\subsubsection{6. Blinding}\label{blinding}}

\begin{itemize}
\tightlist
\item
  No blinding is involved in this study.
\end{itemize}

\hypertarget{is-there-any-additional-blinding-in-this-study}{%
\subsubsection{7. Is there any additional blinding in this
study?}\label{is-there-any-additional-blinding-in-this-study}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{study-design}{%
\subsubsection{8. Study Design}\label{study-design}}

Describe your study design here.

\hypertarget{randomization}{%
\subsubsection{9. Randomization}\label{randomization}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{sampling-plan}{%
\subsection{Sampling Plan}\label{sampling-plan}}

\hypertarget{existing-data}{%
\subsubsection{10. Existing data}\label{existing-data}}

\begin{itemize}
\tightlist
\item
  Registration following analysis of the data: As of the date of
  submission, you have accessed and analyzed some of the data relevant
  to the research plan. This includes preliminary analysis of variables,
  calculation of descriptive statistics, and observation of data
  distributions. Please see cos.io/prereg for more information.
\end{itemize}

\hypertarget{explanation-of-existing-data}{%
\subsubsection{11. Explanation of existing
data}\label{explanation-of-existing-data}}

Describe any existing data or state not applicable.

\hypertarget{data-collection-procedures}{%
\subsubsection{12. Data collection
procedures}\label{data-collection-procedures}}

Enter a description of your data collection procedures here.

\hypertarget{sample-size}{%
\subsubsection{13. Sample size}\label{sample-size}}

Enter a description of your sample size here.

\hypertarget{sample-size-rationale}{%
\subsubsection{14. Sample size rationale}\label{sample-size-rationale}}

Enter your sample size rationale here.

\hypertarget{stopping-rule}{%
\subsubsection{15. Stopping rule}\label{stopping-rule}}

Enter your stopping rule here or state not applicable.

\hypertarget{variables}{%
\subsection{Variables}\label{variables}}

\hypertarget{manipulated-variables}{%
\subsubsection{16. Manipulated variables}\label{manipulated-variables}}

Describe manipulated variables here or state not applicable.

\hypertarget{measured-variables}{%
\subsubsection{17. Measured variables}\label{measured-variables}}

Describe your measured variables here.

\hypertarget{indices}{%
\subsubsection{18. Indices}\label{indices}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{analysis-plan}{%
\subsection{Analysis Plan}\label{analysis-plan}}

\hypertarget{statistical-models}{%
\subsubsection{19. Statistical models}\label{statistical-models}}

Describe your planned statistical model(s) here.

\hypertarget{transformations}{%
\subsubsection{20. Transformations}\label{transformations}}

Describe any transformations here or state not applicable.

\hypertarget{inference-criteria}{%
\subsubsection{21. Inference criteria}\label{inference-criteria}}

Describe your inference criteria here or state not applicable.

\hypertarget{data-exclusion}{%
\subsubsection{22. Data exclusion}\label{data-exclusion}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{missing-data}{%
\subsubsection{23. Missing data}\label{missing-data}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{exploratory-analysis}{%
\subsubsection{24. Exploratory analysis}\label{exploratory-analysis}}

Enter any plans for exploratory data analysis here or state not
applicable.

\hypertarget{other}{%
\subsection{Other}\label{other}}

\hypertarget{other-1}{%
\subsubsection{25. Other}\label{other-1}}

\begin{itemize}
\tightlist
\item
  Not applicable.
\end{itemize}

\hypertarget{references}{%
\subsection{References}\label{references}}

Enter any references used throughout the text here.

\end{document}